Insights Into Computational Reproducibility in Plant Pathology and a Way Forward

### 2021-06-10 analysis/paper/paper.Rmd
edits and comments in the manuscript

# Introduction

Modern plant pathological research has many facets given the array of disciplines and sub-disciplines currently involved.
Collectively, they contribute to increasing our basic and applied knowledge of several aspects of pathogen biology and disease development to ultimately improve plant disease management.
Scientific research in the field varies from the purely observational or descriptive nature to inferential, based on experimental or simulation-derived data sets.
Whatever the case, research findings are verifiable based on how much of the research materials, processes and outcomes are made available beyond what is reported in the scientific article and the ability of others to make use of the methods and results.
These research findings include biological materials (isolates or strains), nucleic/protein sequences, experimental and simulated raw data annotations, drawings and photographs and statistical analysis code among other materials and data generated as a course of conducting research.

@@

While open science and reproducibility are not the same, they are linked.
By proactively practising open science, the work becomes more reproducible through the availability of data and code.
That is, open science leads to reproducibility and replicability.

<!--# not sure we need a separate section for definitions, but presence continuously in the text -->

## Definitions

For us to easily discuss the topic, we first must define what we mean so that we may clearly communicate.
Many of the terms used in this area have varying definitions that may or may not agree with each other.
For instance, reproducible research is recently highlighted by many authors [@Baker2016a; @Iqbal2016; @Nature2016; @Patil2016; @Weissgerber2016; @Brunsdon2015; @Sweedler2015; @Fitzjohn2014; @Ioannidis2014; @Fidler2013; @Stodden2013; ]

<!--# These references above are all prior 2017. We may want to update the list? -->

@@

In an effort to understand where as a discipline we stand as plant pathologists in regards to open science and reproducible research, we surveyed a broad selection of articles to represent a broad swathe of publications to evaluate our status.

@@

# Materials and methods

@@

@@

    -   1 - Uses expensive proprietary software that only institutions would typically purchase (*i.e.*, ArcGIS standard is \$7000)

    -   2 - Uses proprietary software that most individuals can afford (*e.g.*, Excel, Matlab (\$500))

    -   3 - Uses entirely open source and free software (*e.g.*, R, Julia, Python)

    -   NA - No software was used in the research that can be determined as the article is written

@@

    -   1 - Uses expensive proprietary software that only institutions would typically purchase (*e.g.*, ArcGIS standard is \$7000US)

@@

    -   NA - No software was used in the research that can be determined as the article is written

@@

    -   3 - Freely available online to anonymous users for foreseeable future (*e.g.*, archived using Zenodo, dataverse or university library or some other proper archiving system)

@@

# Results

Forty-eight articles scored 0% out of a possible 100% for reproducibility.
The mean value was 31.2% and the median value was 33.3%.
One article scored 100% or was classed as fully reproducible.

No journals surveyed were completely closed, all offered at least an option for open access but some were completely open.

The majority of the articles were classed as fundamental, 160, with the remaining 138 classed as applied.

# Discussion

At the second level, Lehner et al. [-@Lehner2017] used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean tied (<https://emdelponte.github.io/paper-white-mold-meta-analysis/>).
The website clearly demonstrates the analysis for readers and uses R so that anyone can easily replicate the work.
Likewise, Duku et al. [-@Duku2016] provide scripts for models, data and code for graphs, via a website hosted by GitHub, <http://adamhsparks.github.io/MICCORDEA/>, that can be used to replicate their study modelling the effects of climate change on rice bacterial blight and rice leaf blast in Tanzania.

The fourth level includes packaging the analysis as a software package, *e.g.*, R Julia or Python package at a minimum, with the use of continuous integration and possibly a software container like Docker as Khaliq et al. [@Khaliq2020] have done.
Their research compendium provides a Docker instance, DOI and full R package that enable readers to fully replicate their analysis of *Ascochyta rabiei* conida dispersal in chickpea using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a].
When this level is employed, tools such as Binder, <https://mybinder.org>, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks as Mioroni et al. used [-@Miorini2018, [-@Miorini2019]].
Kamvar et al. [-@Kamvar2015] took a slightly different approach by including all files necessary for the analysis and most ouptut files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

To make your data the most widely usable, ensure that it has a persistent identifier.
A digital object identifier (DOI) is the most common (<https://www.doi.org/>) but handle is also an option (<https://handle.net/>).
There are different options for generating a DOI for your data and other materials.
FigShare, Zenodo and OSF all offer long-term archival along with a service to generate a DOI for your materials.
The use of a persistent identifier works to ensure that even if the data are moved, they can still be located using that unique identifier.
For more on FAIR data, visit Go-Fair <https://www.go-fair.org/fair-principles/>.

>>>>

Modern plant pathological research has many facets given the array of disciplines and sub-disciplines currently involved.
Collectively, they contribute to increasing our basic and applied knowledge of several aspects of pathogen biology and disease development to ultimately improve plant disease management.
Scientific research in the field varies from the purely observational or descriptive nature to inferential, based on experimental or simulation-derived data sets.
Whatever the case, research findings are verifiable based on how much of the research materials, processes and outcomes are made available beyond what is reported in the scientific article and the ability of others to make use of the methods and results.
These research findings include biological materials (host and pathogen genotypes), nucleic/protein sequences, experimental and simulated raw data annotations, drawings and photographs and statistical analysis code among other materials and data generated during the course of the research.

@@

## What is reproducibility?

Reproducibility is one component under the umbrella of open science.
By proactively practising open science, the work increases the chance to become more reproducible through the availability of data and code.
That is, open science leads to reproducibility and replicability.

<!--# not sure we need a separate section for definitions, but present continuously in the text -->

For us to easily discuss the topic, we first must define what we mean so that we may clearly communicate.
Many of the terms used in this area have varying definitions that may or may not agree with each other.
For instance, reproducible research is recently highlighted by many authors[@preeyanon2018; @Wallach2018] [@Baker2016a; @Iqbal2016; @Nature2016; @Patil2016; @Weissgerber2016; @Brunsdon2015; @Sweedler2015; @Fitzjohn2014; @Ioannidis2014; @Fidler2013; @Stodden2013; ]

<!--# These references above are all prior 2017. We may want to update the list? I've added two using insert citation tool of the visual editor rmarkdown -->

@@

@@

## Where do we stand in regards to reproducibility?

In an effort to understand where as a discipline we stand as plant pathologists in regards to open science and reproducible research, we surveyed a broad selection of articles to represent a broad swathe of publications to evaluate our status.
We randomly selected 300 articles published from 2012 to 2018 across 21 journals (Table \@ref(tab:journals)) and assigned them scores reflecting their openness and reproducibility.
The journals represented both fundamental and applied areas as well as pathogen or other group-specific publications (See appendix for details on the methodology).

Forty-eight articles scored 0% out of a possible 100% for reproducibility.
The mean value was 31.2% and the median value was 33.3%.
One article scored 100% or was classed as fully reproducible.
No journals surveyed were completely closed, all offered at least an option for open access but some were completely open.The majority of the articles were classed as fundamental, 160, with the remaining 138 classed as applied.

## Are there good examples in the field?

At the second level, Lehner et al. [-@Lehner2017] used GitHub to host a code repository of their research compendium website with data and a reproducible report that explains in detail all steps of the analysis and the R code for conducting a meta-analysis for assessing heterogeneity in the relationship between white mold incidence and soybean yield and between incidence and soybean tied ([<https://emdelponte.github.io/paper-white-mold-meta-analysis/>](https://emdelponte.github.io/paper-white-mold-meta-analysis/){.uri}). The website clearly demonstrates the analysis for readers and uses R so that anyone can easily replicate the work. Likewise, Duku et al. [-@Duku2016] provide scripts for models, data and code for graphs, via a website hosted by GitHub, [<http://adamhsparks.github.io/MICCORDEA/>](http://adamhsparks.github.io/MICCORDEA/){.uri}, that can be used to replicate their study modelling the effects of climate change on rice bacterial blight and rice leaf blast in Tanzania.

The fourth level includes packaging the analysis as a software package, *e.g.*, R Julia or Python package at a minimum, with the use of continuous integration and possibly a software container like Docker as Khaliq et al. [@Khaliq2020] have done. Their research compendium provides a Docker instance, DOI and full R package that enable readers to fully replicate their analysis of *Ascochyta rabiei* conida dispersal in chickpea using the data collected and stepping through other points where weather data were investigated and various models were fit before deciding on the best fit and recreate any figures as published in the article [@Khaliq2020a]. When this level is employed, tools such as Binder, [<https://mybinder.org>](https://mybinder.org){.uri}, can be used that allow readers and reviewers to launch an interactive session in their web browser and interact with the data and rerun the analysis in an RStudio instance or Jupyter notebooks as Mioroni et al. used [-@Miorini2018, [-@Miorini2019]]. Kamvar et al. [-@Kamvar2015] took a slightly different approach by including all files necessary for the analysis and most ouptut files in a repository [@Kamvar2014] that also included an installable R package that was used for the original analysis.
While there are many other methods, these two approaches illustrate some of the best-practices where the data and other files were deposited in repositories with DOIs and reproducibility issues were addressed by using R packages to handle dependencies and other versioning issues making the work more portable.

## Best practices to enhance reproducible research

To make your data the most widely usable, ensure that it has a persistent identifier.
A digital object identifier (DOI) is the most common ([<https://www.doi.org/>](https://www.doi.org/){.uri}) but handle is also an option ([<https://handle.net/>](https://handle.net/){.uri}).
There are different options for generating a DOI for your data and other materials.
FigShare, Zenodo and OSF all offer long-term archival along with a service to generate a DOI for your materials.
The use of a persistent identifier works to ensure that even if the data are moved, they can still be located using that unique identifier.
For more on FAIR data, visit Go-Fair [<https://www.go-fair.org/fair-principles/>](https://www.go-fair.org/fair-principles/){.uri}.

# Appendix

@@

<!--# I would prefer a table for the scores, but if this is to be shown in a appendix, it could stay as it is -->

@@

    -   1 - Uses <!--# should we use past tense or just "expensive .." --> expensive proprietary software that only institutions would typically purchase

    -   <!--# suggest to remove the average prices -->

    -   2 - Proprietary software that most individuals can afford (*e.g.*, Excel)

    -   3 - Entirely open source and free software (*e.g.*, R, Julia, Python)

    -   NA - No software was used in the research that can be determined

@@

    -   1 - Expensive proprietary software that only institutions would typically purchase

@@

    -   NA - No software was used in the research that can be determined

@@

    -   3 - Freely available online to anonymous users for foreseeable future (*e.g.*, archived using Zenodo, dataverse or University library or some other proper archiving system)

@@
